# This script is the reference script for figure 1
```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
```

# Load the data
```{r}
# take the processed object list from the processing script
load(file = "r-objects/object.list.rds")
# bring in color data frame
mac_levels <- read.csv("output/releveling_macrophages.csv")
ct_color_vector <- mac_levels$color
names(ct_color_vector) <- mac_levels$group_var

# for the full object umaps
umap.color.vector <- c(
  "#6666CC",
  "#339900",
  "#33CCCC",
  "#990000",
  "#996600",
  "#9966FF",
  "#CC33CC",
  "#FF9999"
)
names(umap.color.vector) <-
  c(
    "Lymphoid",
    "Epithelial",
    "Mesenchyme",
    "MPs",
    "Endothelial",
    "Cycling",
    "Plasma",
    "cDC"
  )
```

# Figure 1B & C: umap plots with myeloid lineage cells colored in red and then the myeloid lineage umap plots
```{r, Figure 1B & C}
  full.umaps <-
    lapply(object.list, function(obj.list) {
      obj.list[["full"]]$cell.group <- ifelse(obj.list[["full"]]$indepth_cellType %in% names(ct_color_vector),
        "MPs",
        ifelse(obj.list[["full"]]$indepth_cellType %in% c("cDC 1", "cDC 2"),
          "cDC",
          obj.list[["full"]]$indepth_cellType
        )
      )
      p <- DimPlot(obj.list[["full"]],
        label = F,
        raster = FALSE,
        group.by = "cell.group",
        reduction = "umap",
        pt.size = 0.1
      ) +
        scale_color_manual(values = umap.color.vector) +
        ggtitle(
          toupper(obj.list[["full"]]$organ[1]),
          ncol(obj.list[["full"]])
        ) +
        BoldTitle() +
        theme(
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.title = element_blank(),
          plot.subtitle = element_text(hjust = 0.5)
        )
      return(p)
    })
  myeloid.umaps <-
    lapply(object.list, function(obj.list) {
      p <- DimPlot(obj.list[["myeloid"]],
        label = F,
        raster = FALSE,
        group.by = "group_var",
        pt.size = 0.1
      ) +
        scale_color_manual(values = ct_color_vector) +
        ggtitle(
          toupper(obj.list[["myeloid"]]$organ[1]),
          ncol(obj.list[["myeloid"]])
        ) +
        BoldTitle() +
        theme(
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.title = element_blank(),
          plot.subtitle = element_text(hjust = 0.5)
        )
      return(p)
    })

  # Merging legends
  legend_1 <- ggpubr::get_legend(full.umaps[[3]])
  legend_2 <- ggpubr::get_legend(myeloid.umaps[[3]])

  # Combining plots
  rm_legend <- function(p) {
    p + theme(legend.position = "none")
  }
  plots_1 <- ggpubr::ggarrange(plotlist = lapply(full.umaps, rm_legend), nrow = 1)
  plots_2 <- ggpubr::ggarrange(plotlist = lapply(myeloid.umaps, rm_legend), nrow = 1)

  # plots + merged legends
  ggpubr::ggarrange(plots_1, legend_1, widths = c(0.84, 0.16))
  ggpubr::ggarrange(plots_2, legend_2, widths = c(0.84, 0.16))

```

# Figure 1D: heatmap the top 10 DGE per cluster and label the conserved genes
```{r, Figure 1D: top10DGE heatmap}
  # calculating myeloid markers
  mye.markers <-
    lapply(object.list, function(obj.list) {
      mye <- obj.list[["myeloid"]] # select the myeloid specific object
      mye <- NormalizeData(mye) # log normalize
      Idents(mye) <- "group_var"
      # find positive markers, downsample for processing time
      M <- FindAllMarkers(mye, 
        only.pos = TRUE,
        max.cells.per.ident = 5000
      )
      return(M)
    })
  names(mye.markers) <- names(object.list)
  mye.markers.bound <- bind_rows(mye.markers, .id = "organ")

  # retaining top genes per cluster
  all.genes <- unlist(
    lapply(
      mye.markers,
      function(M) {
        genes.to.plot <-
          M %>%
          group_by(gene) %>%
          mutate(rank = rank(-avg_log2FC)) %>%
          filter(rank == 1) %>%
          group_by(cluster) %>%
          top_n(n = 10, wt = avg_log2FC) %>%
          pull(gene)
        return(genes.to.plot)
      }
    )
  )

  # Identify the duplicated genes to label in the plot
  genes.to.label <- all.genes[duplicated(all.genes)]

  annot_col_df <-
    data.frame(
      group_var =
        mac_levels %>%
          select(group_var) %>%
          distinct() %>%
          pull(group_var),
      row.names =
        mac_levels %>%
          select(group_var) %>%
          distinct() %>%
          pull(group_var)
    )

  annot_color_list <-
    list("group_var" = mac_levels$color[match(annot_col_df$group_var, mac_levels$group_var)])
  names(annot_color_list$group_var) <- rownames(annot_col_df)

  heatmap.list <-
    lapply(
      names(object.list),
      function(organ) {
        M <- mye.markers[[organ]]
        genes.to.plot <-
          M %>%
          group_by(gene) %>%
          mutate(rank = rank(-avg_log2FC)) %>%
          filter(rank == 1) %>%
          group_by(cluster) %>%
          top_n(n = 10, wt = avg_log2FC) %>%
          pull(gene)
        ave.exp.df <-
          AverageExpression(object.list[[organ]][["myeloid"]],
            assay = "RNA",
            slot = "data",
            group.by = "group_var",
            features = genes.to.plot
          ) # calculate average expression of genes included
        labs.row <- rownames(ave.exp.df$RNA)
        labs.row[!(labs.row %in% genes.to.label)] <- ""
        p <-
          pheatmap::pheatmap(as.matrix(ave.exp.df$RNA),
            labels_row = labs.row,
            cluster_cols = F,
            cluster_rows = F,
            angle_col = 90,
            annotation_col = annot_col_df,
            annotation_colors = annot_color_list,
            show_colnames = FALSE,
            scale = "row",
            # main = paste0(toupper(organ)),
            color = colorRampPalette(c("#0000CC", "#FFFFFF", "#CC0000"))(100),
            border_color = NA,
            fontsize_row = 4,
            annotation_names_col = FALSE,
            annotation_legend = ifelse(organ %in% c("liver", "kidney"), FALSE, TRUE),
            # annotation_legend = FALSE,
            legend = ifelse(organ %in% c("liver", "kidney"), FALSE, TRUE),
            silent = T,
            drop_levels = FALSE,
          )
        return(p[[4]])
      }
    )

  ggpubr::ggarrange(
    plotlist = heatmap.list,
    common.legend = FALSE,
    ncol = 3,
    widths = c(0.28, 0.28, 0.44)
  )
```

# Figure 2A: Adding M1 and M2 module scores and plotting 
```{r, Figure 2A}
  # using the genes from Viola et al
  module.gene.list <- list(
    "M1" = c(
      "TNF", "IL1B", "IL6", "IL12B", "IL23A", "CD80", "CD86", "CIITA", "HLA-DRB1",
      "NOS2", "PFKFB3", "PKM", "CAD", "STAT1", "HIF1A", "JUN", "FOS"
    ),
    "M2" = c(
      "IL10", "TGFB1", "MRC1", "CD36", "IL1RN", "CD163", "ARG1", "SHPK", "STAT6",
      "GATA3", "SOCS1", "PPARG", "TLR1", "TLR8", "APC", "GLUL"
    )
  )
  # score the cells based on their expression of each module
  object.list <-
    lapply(
      object.list,
      function(obj.list) {
        obj <- obj.list[["myeloid"]]
        DefaultAssay(obj) <- "RNA"
        obj <- NormalizeData(obj)
        obj <- AddModuleScore(obj, features = module.gene.list)
        colnames(obj@meta.data)[c(grep("Cluster1", colnames(obj@meta.data)):ncol(obj@meta.data))] <- names(module.gene.list)
        obj.list[["myeloid"]] <- obj
        return(obj.list)
      }
    )

  # plot the results
  p <-
    lapply(object.list, function(obj.list) {
      md <- obj.list[["myeloid"]]@meta.data
      md <- md %>% select(group_var, names(module.gene.list))
      return(md)
    }) %>%
    bind_rows(.id = "organ") %>%
    mutate(
      group_var = toupper(group_var),
      organ = factor(toupper(organ), levels = c("LIVER", "KIDNEY", "LUNG"))
    ) %>%
    reshape2::melt(variable.name = "module.score") %>%
    mutate(module.score = toupper(module.score)) %>%
    ggplot(aes(y = value, x = group_var, fill = group_var)) +
    geom_boxplot(outlier.size = 0.25) +
    theme_classic() +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    facet_grid(organ ~ module.score) +
    scale_fill_manual(values = ct_color_vector, limits = force) +
    ggtitle("Macrophage Polarization Score") +
    ylab("Module Score") +
    theme(
      axis.text = element_text(color = "black"),
      strip.text.x = element_text(face = "bold"),
      strip.text.y = element_text(face = "bold"),
      legend.position = "top",
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks = element_blank()
    )
  p
```

# Figure 2B: Venn diagram of DGE in MT MACS in each organ
```{r, Figure 2B}
# filter the DGE lists
  mt.mac.genes <-
    lapply(
      mye.markers,
      function(M) {
        genes <- M %>%
          filter(
            cluster == "MAC_MT",
            p_val_adj < 0.05,
            avg_log2FC > 0
          ) %>%
          pull(gene)
        return(genes)
      }
    )
  names(mt.mac.genes) <- toupper(names(mt.mac.genes))
  
  ggvenn::ggvenn(mt.mac.genes,
    fill_color = c(
      "#339900",
      "#CC3333",
      "#6699FF"
    )
  )

  # to identify the genes in each slice
  duplicated.genes <- unlist(mt.mac.genes, use.names = FALSE)[duplicated(unlist(mt.mac.genes))]
  organ.summary <- sapply(
    duplicated.genes,
    function(gene) {
      in.lung <- gene %in% mt.mac.genes[["LUNG"]]
      in.liver <- gene %in% mt.mac.genes[["LIVER"]]
      in.kidney <- gene %in% mt.mac.genes[["KIDNEY"]]

      res <- names(mt.mac.genes)[c(in.lung, in.kidney, in.liver)]
      return(res)
    }
  )
  gene.df <-
    data.frame(
      gene = names(organ.summary),
      organs = unlist(lapply(organ.summary, function(vec) {
        paste(vec, collapse = ",")
      })),
      nHits = unlist(lapply(organ.summary, length))
    ) %>%
    arrange(desc(nHits), organs) %>%
    distinct()
```

# Figure 2C: Violin plot of common DGEs and MT genes
```{r, Figure 2C}
  all.genes <- unique(unlist(lapply(object.list, function(obj.list) {
    return(rownames(obj.list[[2]]))
  })))
  mt.genes <- all.genes[grep("^MT1", all.genes)]
  mt.genes <- c(mt.genes, all.genes[grep("^MT2", all.genes)])
  mt.genes <- c(mt.genes, c("MTF1", "MTF2"))
  mt.genes <- mt.genes[-which(mt.genes == "MT1HL1")]
  mt.genes <- c(mt.genes, "CTSL", "SGK1")

  # generating box plots of mt-gene expression in each organ
  mt.plot.list <- lapply(
    object.list,
    function(obj.list) {
      mye <- obj.list[["myeloid"]]

      # filter to include genes present in the object
      genes.to.violin <- mt.genes[mt.genes %in% rownames(mye)]
      genes.to.add.in <- mt.genes[!(mt.genes %in% rownames(mye))]
      mini_matrix <- matrix(data = 0, nrow = length(genes.to.add.in), ncol = ncol(mye))
      # make zero count matrix for genes not included for visualization purposes (i.e. create blank violin plot)
      rownames(mini_matrix) <- genes.to.add.in

      new.counts <- rbind2(mye[["RNA"]]@data, mini_matrix)
      mye[["plot"]] <- CreateAssayObject(data = new.counts)

      p <-
        VlnPlot(mye,
          features = mt.genes,
          stack = TRUE,
          fill.by = "ident",
          assay = "plot",
          group.by = "group_var",
          same.y.lims = TRUE
        ) +
        scale_fill_manual(values = ct_color_vector) +
        theme(
          axis.text.y = element_text(size = 8),
          axis.text.x = element_text(size = 6, vjust = 4),
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(
            hjust = 1,
            vjust = -2.25,
            size = 10
          ),
          plot.margin = unit(c(-5, 5, -2, 2), "pt")
        ) +
        ggtitle(toupper(mye$organ[1])) +
        NoLegend()
      if (mye$organ[1] == "liver") {
        p4 <- p4 + theme(strip.text.x = element_text(angle = 0, size = 8))
      } else {
        p4 <- p4 + theme(strip.text.x = element_blank())
      }
      return(list(p))
    }
  )
  ggpubr::ggarrange(
    plotlist = lapply(mt.plot.list, function(p.list) {
      return(p.list[[1]])
    }),
    ncol = 1, nrow = 3,
    common.legend = TRUE,
    align = "v"
  )
```

# Figure 3B&C: Generation of gene signatures and adding them as module scores
```{r, Figure 3B&C}
  macs.to.test <- unique(names(ct_color_vector))
for(mac.type in macs.to.test){
  gene.sig.list <- list()
  for(organ in c("liver",
                 "kidney",
                 "lung")) {
    mye <- object.list[[organ]][["myeloid"]]
    DefaultAssay(mye) <- "RNA"
    mye <- NormalizeData(mye)
    
    # Finding markers against all cell types
    Idents(mye) <- "group_var"
    # M <- FindAllMarkers(mye,
    #                     max.cells.per.ident = 5000,
    #                     only.pos = TRUE)
    # write.csv(
    #   M,
    #   file = paste0("output/", organ, "_myeloid_markers_no_DC.csv"),
    #   row.names = FALSE
    # )
    
    marker.file.path <- paste0("output/", organ, "_myeloid_markers_no_DC.csv")
    M <- read.csv(marker.file.path)
    
    ct.to.test <- mac.type
    ct.signature <-
      M %>%
      group_by(gene) %>%
      mutate(gene.rank = rank(-avg_log2FC)) %>%
      filter(gene.rank == 1 &
               cluster == ct.to.test &
               avg_log2FC > 0.75 &
               p_val_adj < 0.05)
    ref.df <-
      M %>%
      filter(gene %in% ct.signature$gene,
             p_val_adj < 0.05) %>%
      group_by(gene) %>%
      mutate(gene.rank = rank(-avg_log2FC)) %>%
      filter(gene.rank == 2)
    
    test.df <-
      left_join(ct.signature[, "gene"], ref.df[, c("gene", "cluster")])
    if (sum(is.na(test.df$cluster)) > 0) {
      res.genes <- test.df$gene[which(is.na(test.df$cluster))]
      test.df <- test.df[-which(is.na(test.df$cluster)), ]
    }
    groups.and.genes <-
      test.df %>% group_by(cluster) %>% group_split()
    gene.tests <-
      lapply(groups.and.genes,
             function(test.group.df) {
               test.group <- as.character(test.group.df$cluster[1])
               test.genes <- test.group.df$gene
               mini.M <-
                 FindMarkers(
                   mye,
                   features = test.genes,
                   ident.1 = ct.to.test,
                   logfc.threshold = 0,
                   ident.2 = test.group,
                   group.by = "group_var"
                 )
               sig.genes <-
                 mini.M %>% filter(p_val_adj < 0.1, avg_log2FC > 0)
               sig.genes <- rownames(sig.genes)
               return(sig.genes)
             })
    
    gene.signature <- unlist(gene.tests)
    if (exists("res.genes")) {
      gene.signature <- c(res.genes, gene.signature)
      rm(res.genes)
    }
    message("Found ", length(gene.signature), " genes in signature for ", ct.to.test, " in ", organ)
    gene.sig.list[[organ]] <- gene.signature
  }
  
  # adding the module score to the objects
  object.list <- 
    lapply(object.list, 
           function(obj.list){
             mye <- obj.list[["myeloid"]]
             mye <- AddModuleScore(mye,
                                   features = list("Cluster1" = gene.sig.list[[organ]]),
                                   assay = "RNA"
             )
             colnames(mye@meta.data)[which(colnames(mye@meta.data) == "Cluster1")] <- paste(mac.type)
             obj.list[["myeloid"]] <- mye
             return(obj.list)
           })
  
  plot.min.max <- range(
    unlist(
      lapply(
        object.list, function(obj.list){
          return(obj.list[["myeloid"]][[paste(mac.type)]])
        }
      )
    )
  )
  
  # creating our signature box plots
  sig.plot.list <- 
    lapply(names(gene.sig.list), 
           function(organ){
             if(length(gene.sig.list[[organ]]) == 0){
               return(NULL)
             }
             mye <- object.list[[organ]][["myeloid"]] 
             p <- 
               mye@meta.data %>%
               dplyr::select(group_var, paste(mac.type)) %>%
               mutate(group_var = toupper(group_var)) %>%
               reshape2::melt(variable.name = "module.score") %>%
               mutate(module.score = toupper(module.score)) %>%
               ggplot(aes(y = value, x = group_var, fill = group_var)) +
               geom_boxplot(outlier.size = 0.01) +
               theme_classic() +
               ggtitle(toupper(mye$organ[1]), 
                       paste(
                         paste(gene.sig.list[[organ]][1:11], collapse = " "),
                         paste(gene.sig.list[[organ]][12:length(gene.sig.list[[organ]])], collapse = " "),
                         sep = "\n"
                       )) +
               # geom_vline(xintercept = 1, color = "red", linetype = "dashed") +
               # geom_vline(xintercept = 0, color = "black") +
               facet_grid(~module.score, scales = "free") +
               ylab("Cell Type") +
               xlab("MT Mac Signature Score") +
               scale_fill_manual(values = ct_color_vector) +
               ylim(plot.min.max) +
               theme(axis.text.x = element_text(size = 6, 
                                                angle = 30, 
                                                vjust = 1,
                                                hjust = 1,
                                                color = "black"),
                     axis.text.y = element_text(size = 6, 
                                                # margin = 0,
                                                # vjust = 1,
                                                hjust = 5,
                                                color = "black"),
                     axis.title = element_blank(),
                     axis.ticks = element_blank(),
                     strip.text.x = element_blank(),
                     plot.title = element_text(hjust = 0.5,
                                               vjust = -1.5,
                                               size = 8,
                                               face = "bold"),
                     plot.subtitle = element_text(hjust = 0, 
                                                  size = 6),
                     plot.margin=unit(c(1,1,1,1),"pt")
               ) 
             if(organ != "liver"){
               p <- p + theme(axis.text.y = element_blank(),
                              plot.margin=unit(c(1,-1,1,1),"pt"))
             }
             return(p)
           })
  
  
  legend <- ggpubr::get_legend(sig.plot.list[[length(sig.plot.list)]])
  rm_legend <- function(p){p + theme(legend.position = "none")}
  plots <- ggpubr::ggarrange(plotlist = lapply(sig.plot.list, rm_legend),
                             nrow = 1, 
                             ncol = 3)
  
  print(plots)
  print(ggpubr::ggarrange(plots, legend, widths = c(0.84, 0.16)))
  

  max.length <- max(unlist(lapply(gene.sig.list, length)))
  myeloid.signature.data.frame <- bind_cols(lapply(gene.sig.list, function(vec){
    if(length(vec) < max.length){
      vec <- c(vec, rep("", max.length-length(vec)))
    }
    return(vec)
  }))
  write.csv(myeloid.signature.data.frame, file = paste0("output/myeloid_signature_gene_sets-",mac.type,".csv"), row.names = FALSE)
}
```

# Figure 3D: Generation of gene signatures in the full object
```{r, Figure 3D}
  # creating signatures for MT Macs only against all other cell types (again)
  full.gene.sig.list <- list()
  for (organ in c(
    "liver",
    "kidney",
    "lung"
  )) {
    obj <- object.list[[organ]][["full"]]
    DefaultAssay(obj) <- "RNA"
    obj <- NormalizeData(obj)

    # Finding markers against all cell types
    Idents(obj) <- "indepth_cellType"
    M <- FindAllMarkers(obj,
      max.cells.per.ident = 5000,
      only.pos = TRUE
    )

    ct.to.test <- "MAC_MT"
    ct.signature <-
      M %>%
      group_by(gene) %>%
      mutate(gene.rank = rank(-avg_log2FC)) %>%
      filter(gene.rank == 1 &
        cluster == ct.to.test &
        avg_log2FC > 0.75 &
        p_val_adj < 0.05)
    ref.df <-
      M %>%
      filter(
        gene %in% ct.signature$gene,
        p_val_adj < 0.05
      ) %>%
      group_by(gene) %>%
      mutate(gene.rank = rank(-avg_log2FC)) %>%
      filter(gene.rank == 2)

    test.df <-
      left_join(ct.signature[, "gene"], ref.df[, c("gene", "cluster")])
    if (sum(is.na(test.df$cluster)) > 0) {
      res.genes <- test.df$gene[which(is.na(test.df$cluster))]
      test.df <- test.df[-which(is.na(test.df$cluster)), ]
    }
    groups.and.genes <-
      test.df %>%
      group_by(cluster) %>%
      group_split()
    gene.tests <-
      lapply(
        groups.and.genes,
        function(test.group.df) {
          test.group <- as.character(test.group.df$cluster[1])
          test.genes <- test.group.df$gene
          mini.M <-
            FindMarkers(
              obj,
              features = test.genes,
              ident.1 = ct.to.test,
              logfc.threshold = 0,
              ident.2 = test.group,
              group.by = "indepth_cellType"
            )
          sig.genes <-
            mini.M %>% filter(p_val_adj < 0.1, avg_log2FC > 0)
          sig.genes <- rownames(sig.genes)
          return(sig.genes)
        }
      )

    gene.signature <- unlist(gene.tests)
    if (exists("res.genes")) {
      gene.signature <- c(res.genes, gene.signature)
      rm(res.genes)
    }
    message("Found ", length(gene.signature), " genes in signature for ", ct.to.test, " in ", organ)
    full.gene.sig.list[[organ]] <- gene.signature
  }

  object.list <-
    lapply(
      object.list,
      function(obj.list) {
        obj <- obj.list[["full"]]
        obj <- NormalizeData(obj)
        obj <- AddModuleScore(obj,
          features = list(
            "MAC_MT" =
              full.gene.sig.list[[obj.list[["full"]]$organ[1]]]
          ),
          assay = "RNA"
        )
        colnames(obj@meta.data)[which(colnames(obj@meta.data) == "Cluster1")] <- "MAC_MT"
        obj.list[["full"]] <- obj
        return(obj.list)
      }
    )

  plot.min.max <- range(
    unlist(
      lapply(
        object.list, function(obj.list) {
          return(obj.list[["full"]]$MAC_MT)
        }
      )
    )
  )
  combined.color.vector <- c(
    "#0000CC",
    "#CCCC00",
    "#33CC99",
    "#6666CC",
    "#FF9999",
    "#CC33CC",
    "#3399FF",
    "#660000",
    "#33CCCC",
    "#333300",
    "#FF9999",
    "#339900",
    "#FF9933",
    "#9966FF",
    "#996600",
    "#999999"
  )
  names(combined.color.vector) <- toupper(
    c(
      "MAC_RES",
      "INF_MONO/MAC",
      "MONOCYTE",
      "Lymphoid",
      "cDC 2",
      "Plasma",
      "MAC_RNASE1+",
      "MAC_SPP1+",
      "Mesenchyme",
      "CYCLING",
      "cDC 1",
      "Epithelial",
      "MAC_MT",
      "Cycling",
      "Endothelial",
      "MAC_misc"
    )
  )
  full.sig.plot.list <-
    lapply(
      names(full.gene.sig.list),
      function(organ) {
        obj <- object.list[[organ]][["full"]]
        p <-
          obj@meta.data %>%
          select(indepth_cellType, MAC_MT) %>%
          mutate(indepth_cellType = toupper(indepth_cellType)) %>%
          reshape2::melt(variable.name = "module.score") %>%
          mutate(module.score = toupper(module.score)) %>%
          ggplot(aes(y = value, x = indepth_cellType, fill = indepth_cellType)) +
          geom_boxplot(outlier.size = 0.01) +
          theme_classic() +
          ggtitle(
            toupper(obj$organ[1]),
            paste(
              paste(full.gene.sig.list[[organ]][1:11], collapse = " "),
              paste(full.gene.sig.list[[organ]][12:length(full.gene.sig.list[[organ]])],
                collapse = " "
              ),
              sep = "\n"
            )
          ) +
          facet_grid(~module.score, scales = "free") +
          ylab("Cell Type") +
          xlab("MT Mac Signature Score") +
          scale_fill_manual(values = combined.color.vector) +
          ylim(plot.min.max) +
          theme(
            axis.text.x = element_text(
              size = 6,
              angle = 30,
              vjust = 1,
              hjust = 1,
              color = "black"
            ),
            axis.text.y = element_text(
              size = 6,
              # margin = 0,
              # vjust = 1,
              hjust = 5,
              color = "black"
            ),
            axis.title = element_blank(),
            axis.ticks = element_blank(),
            strip.text.x = element_blank(),
            plot.title = element_text(
              hjust = 0.5,
              vjust = -1.5,
              size = 8,
              face = "bold"
            ),
            plot.subtitle = element_text(
              hjust = 0,
              size = 6
            ),
            plot.margin = unit(c(1, 1, 1, 1), "pt")
          )

        if (organ != "liver") {
          p <- p + theme(
            axis.text.y = element_blank(),
            plot.margin = unit(c(1, -1, 1, 1), "pt")
          )
        }
        return(p)
      }
    )

  legend <- ggpubr::get_legend(full.sig.plot.list[[3]])
  rm_legend <- function(p) {
    p + theme(legend.position = "none")
  }
  plots <- ggpubr::ggarrange(
    plotlist = lapply(full.sig.plot.list, rm_legend),
    nrow = 1, ncol = 3
  )

  plots
  ggpubr::ggarrange(plots, legend, widths = c(0.84, 0.16))
```

# Figure 4: Interrogation of Mulder et al data
```{r, Figure 4}
  object <- readRDS("r-objects/mulder-et-al/2021_MoMac_VERSE.RDS")

  # Figure 4A: annotated UMAP plots
  DimPlot(object, raster = F, group.by = "Clusters") +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.line = element_blank(),
      axis.title = element_blank(),
      plot.subtitle = element_text(hjust = 0.5)
    )
  object$plot.var <- ifelse(object$Clusters == "Macrophage -11", "MT-Mac", "Other")
  
  # Figure 4B: top 20 DGE of the Mac-11 cluster
  M <- FindMarkers(object, ident.1 = "Macrophage -11", group.by = "Clusters", only.pos = T, max.cells.per.ident = 5000)
  M$gene <- rownames(M)
  M %>%
    filter(p_val_adj < 0.01) %>%
    top_n(n = 20, wt = avg_log2FC) %>%
    arrange(avg_log2FC) %>%
    mutate(
      gene = factor(gene, levels = gene),
      color = ifelse(gene %in%
        c("MT2A", "MT1E", "MT1M", "MT1F", "MT1G", "MT1H", "MT1X", "MT1A", "HAMP"),
      "MT",
      "other"
      )
    ) %>%
    ggplot(aes(x = avg_log2FC, y = gene, fill = color)) +
    geom_bar(stat = "identity") +
    ggtitle("Top 20 DEG of Macrophage -11 Cluster") +
    theme_classic() +
    NoLegend() +
    geom_vline(xintercept = 1, lty = 2) +
    theme(
      axis.text = element_text(color = "black"),
      plot.title = element_text(hjust = 0.5, size = 12, face = "bold")
    ) +
    scale_fill_manual(values = c("#CC0000", "gray"))

  # Figure 4C: % MT in each tissue
  object@meta.data %>%
    group_by(Tissue, Clusters, .drop = FALSE) %>%
    summarise(counts = n()) %>%
    mutate(perc = 100 * counts / sum(counts)) %>%
    filter(Clusters == "Macrophage -11") %>%
    ungroup() %>%
    arrange(perc) %>%
    mutate(Tissue = factor(Tissue, levels = Tissue)) %>%
    ggplot(aes(x = perc, fill = Tissue, y = Tissue)) +
    geom_bar(stat = "identity") +
    scale_x_continuous(labels = scales::percent_format(scale = 1)) +
    theme_classic() +
    NoLegend() +
    xlab("% of MonoMac-Verse") +
    theme(axis.text = element_text(color = "black"))

  DefaultAssay(object) <- "RNA"
  # object <- NormalizeData(object)
  # Add Module score of our genes
  sig.gene.df <- read.csv(file = "output/myeloid_signature_gene_sets.csv")
  conserved.genes <- list(names(table(unlist(sig.gene.list))[table(unlist(sig.gene.list)) == 3]))
  object <- AddModuleScore(object, features = conserved.genes, assay = "RNA")

  # Figure 4D: Module score feature plot
  FeaturePlot(object, features = "Cluster1", order = T, raster = F, reduction = "umap", cols = c("yellow", "red")) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.line = element_blank(),
      axis.title = element_blank(),
      plot.subtitle = element_text(hjust = 0.5)
    )

    # Figure 4E: box plot of module score distribution
  object@meta.data %>%
    select(Clusters, Cluster1) %>%
    reshape2::melt(variable.name = "module.score") %>%
    mutate(module.score = toupper(module.score)) %>%
    ggplot(aes(y = value, x = Clusters, fill = Clusters)) +
    geom_boxplot(outlier.size = 0.01) +
    theme_classic() +
    ggtitle("MT Gene Signature Score") +
    ylab("Cell Type") +
    xlab("MT Mac Signature Score") +
    theme(
      axis.text.x = element_text(
        size = 10,
        angle = 90,
        vjust = 1,
        hjust = 1,
        color = "black"
      ),
      axis.text.y = element_text(
        size = 10,
        # margin = 0,
        # vjust = 1,
        hjust = 5,
        color = "black"
      ),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      strip.text.x = element_blank(),
      plot.title = element_text(
        hjust = 0.5,
        vjust = -1.5,
        size = 12,
        face = "bold"
      ),
      plot.subtitle = element_text(
        hjust = 0,
        size = 6
      )
    ) +
    NoLegend()

  # Figure 4F: Violin plot of each gene in the gene signature 
  VlnPlot(object, features = c(conserved.genes[[1]], "Cluster1"), stack = T, group.by = "Clusters", fill.by = "ident") +
    NoLegend() +
    theme(
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(size = 6, vjust = 4),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      plot.title = element_text(
        hjust = 1,
        vjust = -2.25,
        size = 10
      ),
      strip.text.x = element_text(angle = 0, size = 8)
    )
```

# Figure 5: Ligand-receptor analysis and visualization
```{r, Figure 5}
  library(CellChat)
  cc.list <- list()
  df.net.list <- list()

  obj <- object.list[["liver"]][["full"]]
  DefaultAssay(obj) <- "RNA"
  obj <- NormalizeData(obj)

  cellchat <- createCellChat(object = obj, group.by = "indepth_cellType")
  CellChatDB <- CellChatDB.human
  CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling")
  cellchat@DB <- CellChatDB.use
  cellchat <- subsetData(cellchat)

  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)

  cellchat <- computeCommunProb(cellchat)
  cellchat <- computeCommunProbPathway(cellchat)

  df.net <- subsetCommunication(cellchat)

  # viz time
  cellchat <- aggregateNet(cellchat)
  groupSize <- as.numeric(table(cellchat@idents))

  mt.net <- df.net %>% filter(source == "MAC_MT" | target == "MAC_MT")
  mt.ints <- unique(mt.net$interaction_name)
  mt.paths <- unique(mt.net$pathway_name)
 
  cellchat <- cc.list[["liver"]]
  df.net <- df.net.list[["liver"]]

  pathways.to.show <- c("CHEMERIN", "VEGF", "TWEAK")
  cellchat$indepth_cellType <- toupper(cellchat$indepth_cellType)

  # Figure 4B: Interaction strength dot plot
  netVisual_bubble(cellchat, signaling = pathways.to.show, remove.isolate = FALSE)
  
  # Figure 4C: chord signalling plots
  lapply(pathways.to.show, function(pathway) {
    lapply(c("chord", "circle"), function(plot.type) {
      netVisual_individual(cellchat,
        signaling = pathway,
        remove.isolate = F,
        color.use = paste(combined.color.vector[match(toupper(rownames(obj@netP$prob[, , obj@netP$pathways == "VEGF"])), names(combined.color.vector))]),
        pairLR.use = df.net %>% filter(pathway_name == pathway, source == "MAC_MT" | target == "MAC_MT") %>% pull(interaction_name),
        layout = plot.type
      )
    })
  })

  # Figure 4D: Violin plot of interacting genes
  ggpubr::ggarrange(VlnPlot(obj,
    features = c("RARRES2", "CMKLR1", "VEGFB", "FLT1", "TNFSF12", "TNFRSF12A"),
    pt.size = 0,
    stack = T,
    fill.by = "ident",
    group.by = "indepth_cellType"
  ) +
    scale_fill_manual(values = combined.color.vector) +
    theme(
      axis.text.x = element_text(color = "black"),
      strip.text.x = element_text(angle = 0, vjust = 0),
      axis.title = element_blank()
    ) +
    NoLegend(), nrow = 2, heights = c(0.6, 0.4))
```

